---
title: '04 Quiz: Take-Home'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Complete this Quiz in a .Rmd file. Submit a .Rmd and a knitted .html file to GitHub by the due date and time.

## Statement of Integrity
“All work presented is my own, and I have followed all rules for collaboration.” Jessica Tom

## Question
**Use the leaflet package and the us.cities data set from the maps package to make a Zoom-able map of the 50 U.S. capital cities. When you click on a marker for a capital city, both the city name and state as well as the pop (population) should be displayed.**

**Note that, in the us.cities data set, state capitals are given a value of 2 in the capital variable.**

**A second note is that if you cannot figure out how to get two “things” (in this case name and population) to display in Leaflet, you would just lose a single point for the quiz (so don’t spend too much time on that part unless you are inclined to do so).**

```{r}
library(tidyverse)
library(maps)
library(leaflet)

us.cities <- us.cities %>% filter(capital == 2)
leaflet(us.cities) %>%
  setView(lng = -98.583, lat = 39.833, zoom = 4) %>% 
  addTiles() %>% 
  addProviderTiles(providers$Wikimedia) %>% 
  addMarkers(lng = us.cities$long, lat = us.cities$lat,
             clusterOptions = markerClusterOptions(),
             popup = paste("Capital and State:", us.cities$name, "<br>", 
      "Population:", us.cities$pop, "<br>"))
```