---
title: "Data Visualization Midterm: Take-Home Part 2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# Mini-Tutorial 2: Mapping (15 points)

Topics to Cover:

Either mapping with the maps package OR mapping from a shapefile. Many shapefiles can be found at https://www.census.gov/geographies/mapping-files/time-series/geo/carto-boundary-file.html
Make sure to include:

(4 points) An introductory statement with the goal of the section (in your own words) and a brief description of the data set and variables you will be using.

(11 points) R code chunks interspersed with some text descriptions about the code you have written and with text description about each topic. If you are using the maps package, you should use a map other than the U.S. States Map or World Map that we used in class.

## Overview
This section aims at showing you the basic functions of mapping with RStudio's leaflet. We will be using `world.cities`, which documents world cities with a population more than 40,000 people, in the `maps` package. With this data set, we will make an interactive map that shows you the cities' name, country, and population. Thus, the important variables are:

* `namme` - name of the city
* `country.etc` - country that the city is in
* `population` - population of the city

```{r}
library(tidyverse)
library(maps)
```

## Leaflet
First, we need to create a map. To do so, we will use leaflet's map options--I chose Esri.WorldTopoMap for this example. We also need to designate where the world cities' longitude and latitude coordinates are in `addMarkers`. This function also allows you to set the markers for the cities in clusters so the viewer can view how many major cities are in each region. `setView` allows you to adjust where the viewer initially starts on the map. In this example, the map initially hovers over the U.S. 
```{r}
library(leaflet)

leaflet(world.cities) %>%
  setView(lng = -98.583, lat = 39.833, zoom = 4) %>% 
  addTiles() %>% 
  addProviderTiles(providers$Esri.WorldTopoMap) %>% 
  addMarkers(lng = world.cities$long, lat = world.cities$lat,
             clusterOptions = markerClusterOptions())
```

To add the popup label for the city, country, and population, under `addMarkers` add the desired labels. Follow the formatting of the code below under `popup = paste()`, which indicates the name of the label and where from the data set to get the information. 
```{r}
library(leaflet)

leaflet(world.cities) %>%
  setView(lng = -98.583, lat = 39.833, zoom = 4) %>% 
  addTiles() %>% 
  addProviderTiles(providers$Esri.WorldTopoMap) %>% 
  addMarkers(lng = world.cities$long, lat = world.cities$lat,
             clusterOptions = markerClusterOptions(),
             popup = paste("City:", world.cities$name, "<br>",
                           "Country:", world.cities$country.etc, "<br>",
                           "Population:", us.cities$pop, "<br>"))
```
          